---
title: "Statistical Modeling"
output:
  pdf_document:
    includes:
      in_header: header.tex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library("vcd")
```

2.2 Statistics vrs Probability
Probability deals with predicting the likelihood of future events, while statistics involves the analysis of the frequency of past events. 

2.3 A simple example of statistical modeling
Load and clean data from outleirs
```{r}
load(here("Book", "data", "e100.RData"))
e99 = e100[-which.max(e100)]
```

Plot bar-chat since the data is dicrete
```{r}
barplot(table(e99), space = 0.8, col = "chartreuse4")
```
Goodness-of-fit : visual evaluation
The rootogram (Cleveland 1988); which hangs the bars with the observed counts from the theoretical red points. 

```{r rootogram}
gf1 = goodfit( e99, "poisson")
rootogram(gf1, xlab = "", rect_gp = gpar(fill = "chartreuse4"))
```

Question 1
calibrate a barplot from a known Poisson variable using rpois with λ = 0.05 to generate 100 Poisson distributed numbers and draw their rootogram.

```{r}
simp = rpois(100, lambda = 0.05)
gf2 = goodfit(simp, "poisson")
rootogram(gf2, xlab = "")
```