---
title: "Chapt1_walk through"
author: "Abdul-Rahman"
date: "21/09/2019"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries in this Chapter

```{r package}
library(here)
```

#Poisson distribution
How often can 3 mutations could occur under a Poisson model with a rate parameter of 5

Using X as the event (3) and lambda as the rate parameter (5)

```{r}
dpois(x = 3, lambda = 5) 
```

What are the indiviadual probabilities of 0 to 12 mutations with a rate parameter of 5. Visalualize these probabilities with a barplot

```{r}
dpois(x = 0:12, lambda = 5)
barplot(dpois(0:12, 5), names.arg = 0:12, col = "green",xlab = "Number of mutations", ylab = "Probability")
```

Finding the frequencies of categories in a data (e.g bloodgroup)
```{r bloodgroup}
bloodgroup = c("AA","AO","BB","AO","OO","AO","AA","BO","BO",
             "AO","BB","AO","BO","AB","OO","AB","BB","AO","AO")

table(bloodgroup)
```

Creating a factor that can be used by the level function in identifying discrete categories

```{r}
bloodgroupF = factor(bloodgroup)
levels(bloodgroupF)
```

## Adding factors not captured in the 
##Bernoulli trials
#Trail with equal probabilities
The outcome of 15 Bernoulli trials with a probability of success equal to 0.5
```{r}
rbinom(15, prob = 0.5, size = 1)
```
Each trail is independent of the other hence the difference in the answers when run

#Trail with unequal probabilities
```{r}
rbinom(12, prob = 2/3, size = 1)
```

#Binomial random variable
```{r}
set.seed(235569515)
rbinom(1, prob = 0.3, size = 15)
```

#Probability mass distribution displayed in 2 decimal places
```{r}
probabilities = dbinom(0:15, prob = 0.3, size = 15)
round(probabilities, 2)
```

#Bar plot ofthe probabilities
```{r}
barplot(probabilities, names.arg = 0:15, col = "green", xlab="Number of trials", ylab = "Probability")
```


#Trail question
What is the output of the formula for k=3, p=2/3,n=4?
```{r}
dbinom(3, size = 4, prob = 2/3)
x = factorial(4)/factorial(1)/ factorial(3)* (2/3)^3*(1/3)
```

Simulation
Simulate a mutation process along 10,000 positions with a mutation rate of 
5x10^-4 and count the number of mutations. Repeat this many times and plot the distribution with the barplot function

```{r}
rbinom(1, prob = 5e-4, size = 10000)
```

```{r}
simulations = rbinom(n = 300000, prob = 5e-4, size = 10000)
barplot(table(simulations), col = "lavender")
```
##Modelling (generative) epitope detection

False positive rate for ELISA is 1%
Tests protein in 100 independent positions


Task: Verify by simulation that the sum of 50 independent Bernoulli variables with p=0.01 is – to good enough approximation– the same as a Poisson(0.5) random variable.

```{r ELISA array results for 50 patients in the 100 positions}
set.seed(235569515)
sum(rbinom(50, prob = 0.01, size = 1))
dpois(x = 50, lambda = 0.5)
```

```{r}
load(here("Book", "data", "e100.RData"))

barplot(e100, ylim = c(0, 7), width = 0.7, xlim = c(-0.5, 100.5),
        names.arg = seq(along = e100), col = "darkolivegreen")

```


This is  equivalent to 1−P(  X≤6); the so-called cumulative distribution function at 6. The ppois function can be used for this in two methods;


```{r ppois Method 1}
1 - ppois(6, 0.5)
```

```{r ppois Method 2}
ppois(6, 0.5, lower.tail = FALSE)
```

In method two;
If lower.tail = TRUE(default) then probabilities are P[X ≤ x], meaning probability of 6, 5, 4, 3, 2, 1, 0 people == proability of less than 7
If lower.tail = FALSE then probabilities are  P(X > x) so P(X>6) meaning probability of 7,8,9,10,11, ..., n people from 7 to 100 == proability => 7

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()






list1 <- list()#create empty list.
for (i in 1:200000) {
  Q1 <- rbinom(1000,1,0.0001) #randomly enerate 0/1 randomly.
  Q11 <- sum(Q1) #sum positions to count mutations.
  list1[[i]] <- Q11 # add te sims to a list.
}
list11 <- unlist(list1)# convert te list to a vector
list11
barplot(list11,col = "chartreuse4")#plot
library("vcd") #to do a rootogram
qf1 = goodfit( list11)
rootogram(qf1, xlab = "", rect_gp = gpar(fill = "chartreuse4"))